---
title: "Assignment 1"
date: "`r Sys.Date()`"
output:  pdf_document
---

```{r DocSetup, include = FALSE}
require(knitr)
opts_chunk$set(echo = FALSE, warning = FALSE)
opts_chunk$set(fig.width = 8, fig.height = 6, fig.align = "center")
```



```{r }
################################################################################
#
# Question 1
#
################################################################################

```{r Q1Setup, include = FALSE}
source("Question1/1setup.R")
```


## Question 1: Inference for the binomial parameter:
(a) Develop an R function to calculate HPD intervals for data $(x,n)$, given a \emph{beta(a,b)} prior.

```{r Question1a, eval = FALSE}
solve.HPD.beta = function(h, y, n, a, b, p){
  apost <<- y + a
  bpost <<- n - y + b
  post_mode = (y + a - 1)/(n + a + b - 2)
  lt = uniroot(f=function(x){ dbeta(x,apost, bpost) - h},
               lower=0, upper=mode)$root
  ut = uniroot(f=function(x){ dbeta(x,apost, bpost) - h},
               lower=mode, upper=1)$root
  coverage = pbeta(ut, apost, bpost) - pbeta(lt, apost, bpost)
  hpdval = abs(p-coverage)
  results <<- data.frame(lt,ut,coverage,h)
  return(hpdval)}

  # if (plot) {
  #   th = seq(0, 1, length=5000)
  #   plot(th, dbeta(th, apost, bpost),
  #        t="l", lty=1,xlab=expression(theta),
  #        ylab="posterior Density", ...)
  #   abline(h=h)
  #   segments(ut,0,ut,dbeta(ut,apost,bpost))
  #   segments(lt,0,lt,dbeta(lt,apost,bpost))
  #   title(bquote(paste("p(", .(round(lt, 5))," < ", theta, " < ",
  #                      .(round(ut,5)), " | " , y, ") = ",
  #                      .(round(coverage, 5)), ")")))


y=23; n=80; a=2; b=17; p=0.95
upper <- max(dbeta(seq(0, 1, length=500),y + a,n-y+b))
interval <- seq(0,upper,by = 0.00001)

h <- optimize(solve.HPD.beta,
         interval = interval,
         lower = min(interval),
         tol = .Machine$double.eps,
         y=y,
         n=n,
         a=a,
         b=b,
         p=p)$minimum

  th = seq(0, 1, length=5000)
  plot(th, dbeta(th, apost, bpost),
       t="l", lty=1,xlab=expression(theta),
       ylab="posterior Density")
  abline(h=h)
  segments(ut,0,ut,dbeta(ut,apost,bpost))
  segments(lt,0,lt,dbeta(lt,apost,bpost))
  title(bquote(paste("p(", .(round(lt, 5))," < ", theta, " < ",
                     .(round(ut,5)), " | " , y, ") = ",
                     .(round(coverage, 5)), ")")))
```


Running the above code gives the following output

```{r Q1aplot, eval = FALSE}
source("Question1/1a.R")
```


(b) Reproduce Agresti \& Coull's (1998) Figure 4 $(n = 10)$, and replicate for the Score and Bayes-Laplace \& Jeffreys HPD intervals

```{r Question1bData, echo = FALSE, cache = TRUE}
n <- 10
a <- 0.05
p <- seq(0.0001,0.9999,1/1000)
z <- abs(qnorm(.5*a,0,1))
source("Question1/1bData.R")
```



```{r Question1bChart, echo = FALSE}
source("Question1/1bChart.R")
q1bchart
```


(c) Repeat (b) for $n = 50$.

```{r Question1cData, echo = FALSE, cache = TRUE}
n <- 50
a <- 0.05
p <- seq(0.0001,0.9999,1/1000)
z <- abs(qnorm(.5*a,0,1))
source("Question1/1cData.R")
```



```{r Question1cChart, echo = FALSE}
source("Question1/1cChart.R")
q1cchart
```


(d) Compare the minimum coverage of the six graphs at (c)

```{r Question1d}
source("Question1/1d.R")
```


(e) The adjusted Wald interval appears to perform well with respect to frequentist coverage, if close to nominal combined with reasonable minimum coverage is aimed for.
From a Bayesian point of view, performance of individual intervals is just as, if not more, important. 
Given $x = 0$, compare the adjusted Wald interval with the exact \& Score intervals (all two-sided), and with the Bayes-Laplace \& Jeffreys HPD intervals, for a range of values of $n$ and $\alpha$, comment on its limitations, and give an appropriate graphical illustration.

```{r Question1e}
source("Question1/1e.R")
```


\newpage


```{r }
################################################################################
#
# Question 2
#
################################################################################

```{r Q2Setup, include = FALSE}
rm(list = ls())
source("Question2/2setup.R")
opts_chunk$set(warning = FALSE)
opts_chunk$set(fig.width = 5, fig.height = 3, fig.align = "center")
```


## Question 2: Inference for the Cauchy parameter:

(a) Develop an R function to find percentiles of a (general) Cauchy posterior as discussed by Jaynes (1976, Example 6) and Box \& Tiao (1973, p.64), to be used for the examples below.

```{r Question2a, eval = FALSE}
source("Question2/2a.R")
```


(b) Consider Jaynes' example of $n = 2$ observations $(3, 5)$: plot the posterior and calculate the 90\% central credible interval.
Explain why it is quite different from the confidence interval derived by Jaynes (p.202).

```{r Question2b}
source("Question2/2b.R")
Q2bChart
```


(c) Consider Box \& Tiao's example of $n = 5$ observations $(11.4, 7.3, 9.8, 13.7, 10.6)$: plot the posterior and calculate 95\% central and HPD credible intervals and check $Pr[\theta < 11.5]$ given by Box \& Tiao.

```{r Question2c}
source("Question2/2c.R")
Q2cChart
```


(d) Consider Berger's (1985, p.141) example of $n = 5$ observations $(4.0, 5.5, 7.5, 4.5, 3.0)$: calculate 95\% central and HPD credible intervals, with and without Berger's restriction $(\theta > 0)$.

```{r Question2d}
source("Question2/2d.R")
Q2dChart
```


(e) Clearly, Berger's restriction $(\theta > 0)$ will sometimes lead to a posterior quite different from the unrestricted posterior.
Plot this restricted posterior for the hypothetical negative version of Berger's example: i.e. $(-4.0, -5.5, -7.5, -4.5, -3.0)$, and calculate the 95\% HPD interval.

```{r Question2e}
source("Question2/2e.R")
Q2eChart
```


