Firstly let's import the data

```{r, include = FALSE}
FMDData <- readRDS("~/R/STAT4102-AdvancedBayesianStatistics/Assignment3/Question1/Data/FMDData.rds")

FMDData2 <- FMDData %>%
  mutate(StCattle = (Cattle-mean(Cattle))/sd(Cattle)) %>%
  dplyr::select(-c(Province,Cattle))

FMDData3 <- FMDData %>%
  mutate(StCattle = (Cattle-mean(Cattle))/sd(Cattle)) %>%
  dplyr::select(-Cattle)


fm_pois <- glm(FMD1998 ~ .,data = FMDData2, family = poisson)
fm_poisInteraction <- glm(FMD1998 ~ EasternTurkey*StCattle, data = FMDData2, family = poisson)
fm_poisCattle <- glm(FMD1998 ~ StCattle, data = FMDData2,  family = poisson)
```


```{r}
kable(head(FMDData), caption = "Raw FMD dataset")
```

Let's remove the unnecessary variables (**Province**) and standardise the **Cattle** variable

```{r}
kable(head(FMDData2), caption = "FMD dataset with standardised Cattle values")
```

Next, we fit a basic GLM using the poisson family option in R: *fm_pois <- glm(FMD1998 ~ .,data = FMDData2, family = poisson)*

```{r}
kable(tidy(fm_pois), caption = "Model Output")
kable(glance(fm_pois), caption = "Model Summary")
```

It seems as if the only significant covariate is the (standardised) **Cattle** variable; the **EasternTurkey** region variable does not return a signifcant ($\alpha \leq 0.05$) p value.

Just to check that we aren't missing anything, let's also look at the model summaries for the model including an interaction term **StCattle x EasternTurkey** and the model with only **StCattle** as a covariate.

```{r}
kable(glance(fm_poisInteraction), caption = "Model (with Interaction term) Summary")
kable(glance(fm_poisCattle), caption = "Model (only StCattle) Summary")
```

The original model and the model with only the **StCattle** covariate are -- unsurprisingly -- the best performing models, given that we found the **EasternTurkey** covariate to be insignificant at the beginnning. Whilst the origial model proves the superior one when comparing **deviance, AIC, BIC** and **logLik** scores, the **StCattle** model does have an extra degree of freedom for the residuals, which might be desirable.
