---
title: "Assignment 2"
date: "`r Sys.Date()`"
output:  
  pdf_document:
geometry: margin=1.5cm

---

```{r DocSetup, include = FALSE}
require(knitr)
opts_chunk$set(warning = FALSE, cache = FALSE, echo = FALSE)
opts_chunk$set(fig.height = 3, fig.width = 7,fig.align = "center")
```



```{r Q1Setup, include = FALSE}
source("setup.R")
```


## Question 1: 

Consider data on Foot-and-Mouth disease (FMDdata.xls) with the foal of assessing the difference in incidence between the eastern and western regions of Turkey, based on data from 1998:

(a) Apply classical Poisson regression to the FMD counts, with region and size of cattle population (standardised) as covariates. Check model diagnostics. [10 marks]

```{r child = "Question1/1a.Rmd"}
```


(b) Repeat (a), using a Bayesian approach, and compare results. [10 marks]

```{r child = "Question1/1b.Rmd"}
```

(c) Use a Bayesian zero-inflated Poisson (ZIP) model instead. [10 marks]

```{r child = "Question1/1c.Rmd"}
```


(d) Compare the above two Bayesian models and, for your model of choice, report on any difference in mean FMD occurrence for eastern vs western Turkey. [10 marks]

\newpage

## Question 2:

Consider the AR(1) model (see Phillips, 1991 and Berger & Yang, 1994):

Assuming that $| \rho | \leq 1$, compare posteriors (and intervals) arising from the uniform prior and the reference prior for some generated data sets, based on \emph{e.g} $\rho = 0.1$ and $\rho =0.99$, and sample sizes $n = 50$ and $n = 200$. Also compare with confidence intervals.

```{r child = "Question2/2.Rmd"}
```

## References 
```{r child = "Ref.Rmd"}
```


