---
title: "2a"
author: "Benjamin Moran"
date: "8 May 2017"
output: html_document
---

```{r setup, include = FALSE}
rm(list = ls())
library(knitr)
library(tidyverse)
library(lme4)
library(broom)
opts_chunk$set(warning = FALSE, echo = TRUE, cache = TRUE)
opts_chunk$set(fig.width = 8, fig.height = 6, fig.align = "center")
data("Dyestuff2")
```

(a) Use e.g. SAS (PROC VARCOMP) to perform a classical analysis of the data in Table 5.1.4 of Box & Tiao (1973), based on finding point estimates only.

I'll be using R. First, let's confirm the results reported by Box & Tiao (1973), using frequentist methods.

```{r anova}
library(lme4)
data("Dyestuff2")
D2aov <- aov(Yield ~ Batch, Dyestuff2)
(D2summary <- summary(D2aov))
(sigma2_with <- D2summary[[1]]$`Mean Sq`[2])
(sigma2_btw <- (D2summary[[1]]$`Mean Sq`[1] - sigma2_with)/D2summary[[1]]$Df[1])
```

These results match what was reported by Box & Tiao (1973). Next, we can attempt to fit a linear model using the REML method.

```{r}
(D2REML <- lmer(Yield ~ 1 + (1|Batch), Dyestuff2, REML=TRUE))
D2REML <- broom::tidy(D2REML)
(sigma2_with <- D2REML[which(D2REML$group == "Residual"),"estimate"]^2)
(sigma2_with <- D2REML[which(D2REML$group == "Batch"),"estimate"]^2)
```

We can see here that the between Batch variance has been set to $0$ automatically by the REML method, because it uses
